---
title: "Routing Problem"
output: html_notebook
---

## Introduction

Here, let's try to solve routing problem that is given the list of location and distance between each pair of location, what is the shortest path to visit all location exactly one and return to the origin location.

In general this routing problem have several basic definition including:
- Set of vertices (locations)
- Set of edge (connection between locations)

Edge has an associate distance d>0 which could represent travel time, distance, monetary cost associated with the edge

For example, lets continue our case earlier from warehouse location problem, after we find 4 location to build our warehouse, we want to know what is the distribution route we gonna use that minimize the distance for the delivery to all the warehouse.


## Restrusturing dataset and set parameter

```{r, echo=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
```

```{r}
# Define the number of warehouse
n = 4
```

```{r}
warehouse_loc_r = warehouse_locations %>%
  filter(id %in% c(7,8,15,18))

warehouse_loc_r$id[warehouse_loc_r$id==7]<-1
warehouse_loc_r$id[warehouse_loc_r$id==8]<-2
warehouse_loc_r$id[warehouse_loc_r$id==15]<-3
warehouse_loc_r$id[warehouse_loc_r$id==18]<-4

head(warehouse_loc_r)
```

```{r}
ggplot(warehouse_loc_r, aes(x,y)) +
  geom_point()
```

```{r}
distance <- as.matrix(stats::dist(select(warehouse_loc_r, x, y), diag = TRUE, upper = TRUE))

dist_fun <- function(i, j) {
  vapply(seq_along(i), function(k) distance[i[k], j[k]], numeric(1L))
}

distance
```

## Model Formulation

```{r}
library(ompr)
```

```{r}
model <- MIPModel() %>%
  # we create a variable that is 1 iff we travel from city i to j
  add_variable(x[i, j], i = 1:n, j = 1:n, type = "integer", lb = 0, ub = 1) %>%
  
  # a helper variable for the MTZ formulation of the tsp
  add_variable(u[i], i = 1:n, lb = 1, ub = n) %>% 
  
  # minimize travel distance
  set_objective(sum_expr(dist_fun(i, j) * x[i, j], i = 1:n, j = 1:n), "min") %>%
  
  # you cannot go to the same city
  set_bounds(x[i, i], ub = 0, i = 1:n) %>%
  
  # leave each city
  add_constraint(sum_expr(x[i, j], j = 1:n) == 1, i = 1:n) %>%
  #
  # visit each city
  add_constraint(sum_expr(x[i, j], i = 1:n) == 1, j = 1:n) %>%
  
  # ensure no subtours (arc constraints)
  add_constraint(u[i] >= 2, i = 2:n) %>% 
  add_constraint(u[i] - u[j] + 1 <= (n - 1) * (1 - x[i, j]), i = 2:n, j = 2:n)

model
```

## Result

```{r}
library(ompr.roi)
library(ROI.plugin.glpk)
```

```{r}
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))
```

```{r}
solution <- get_solution(result, x[i, j]) %>% 
  filter(value > 0) 

kable(solution)
```

```{r}
paths <- select(solution, i, j) %>% 
  rename(from = i, to = j) %>% 
  mutate(trip_id = row_number()) %>% 
  tidyr::gather(property, idx_val, from:to) %>% 
  mutate(idx_val = as.integer(idx_val)) %>% 
  inner_join(warehouse_loc_r, by = c("idx_val" = "id"))

kable(arrange(paths, trip_id))
```


```{r}
ggplot(warehouse_loc_r, aes(x, y)) + 
  geom_point() + 
  geom_line(data = paths, aes(group = trip_id)) + 
  ggtitle(paste0("Optimal route with cost: ", round(objective_value(result), 2)))
```


